---
title: "Reproducible Research: Peer Assessment 1"
author: "jenwong-hk"
date: "April 18, 2018"
output: 
  html_document: 
    keep_md: true
---

# What is mean total number of steps taken per day?
download, unzip & import

```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile="C:/Users/HP/Documents/reproductiveResearch/activity.zip")

library(plyr)
zipF <- list.files(path="C:/Users/HP/Documents/reproductiveResearch", pattern = "*.zip", full.names = TRUE)
ldply(.data = zipF, .fun=unzip, exdir = "C:/Users/HP/Documents/reproductiveResearch")

activity <- read.csv("C:/Users/HP/Documents/reproductiveResearch/activity.csv")
```

#Calculate the total number of steps taken per day - ignore missing data
```{r}
library(lubridate)
activity$date <- ymd(activity$date)
data1 <- aggregate(activity$steps, by=list(date=activity$date), FUN=sum, na.rm=FALSE)
```

Make a histogram of the total number of steps taken each day
```{r}
plot(data1, ylab="steps", type="h", lwd=3, main="Total Number of Steps taken per day")
```

Calculate and report the mean and median of the total number of steps taken per day
```{r}
library(dplyr)
data2 <- aggregate(activity$steps, by=list(date=activity$date), FUN=mean, na.rm=FALSE)
temp <- c("date", "mean-number-of-steps-per-day")
colnames(data2) <- temp
print(data2)

data3 <- aggregate(activity$steps, by=list(date=activity$date), FUN=median, na.rm=FALSE)
temp <- c("date", "median-number-of-steps-per-day")
colnames(data3) <- temp
print(data3)
```

What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)
```{r}
data4 <- aggregate(steps ~ interval, data=activity, mean)
plot(data4, type="l", lwd=2, main="Average steps on 5-minute interval accross all days")
```

Which 5-minute interval, on average across all the days in the dataset, 
the maximum number of steps?
```{r}
temp <- subset(data4, steps==max(data4$steps))
temp$interval
```

# Imputing missing values

Calculate and report the total number of rows with NAs
```{r}
good <- complete.cases(activity)
sum(good==F)
```

# Filling in all of the missing values in the dataset 
with the mean for that 5-minute interval, new dataset generated
```{r}
library(dplyr)
y <- data.frame()
i=1
for (i in 1:288){
        data4_steps <- data4$steps[i]
        data4_interval<- data4$interval[i]
        
        x <- filter(activity, activity$interval==data4_interval)
        x[is.na(x)] <- data4_steps
        y <- rbind(x, y)
}

y <- y[order(y$date, y$interval),]
```

histogram of the total number of steps taken each day on new dataset
```{r}
library(lubridate)
y$date <- ymd(y$date)

data5 <- aggregate(steps ~ date, data=y, sum)
plot(data5, type="h", lwd=2, main=" Total number steps each day adj missing data")
```

mean & median total number of steps taken each day
```{r}
aggregate(steps ~ date, data=y, mean)
aggregate(steps ~ date, data=y, median)
```

#What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
data7 <- cbind(data1,data5)
data7 <- data7[,-3]
temp <- c("date","steps.with.NAs", "steps.adj")
colnames(data7) <- temp

with(data7, plot(date, steps.with.NAs, type="l", lwd=2, ylab="steps total per day",
                 main="Impact-Input missing data on total daily number steps"))

with(data7, lines(date, steps.adj, lty=2, lwd=2, col=rgb(1, 0, 0, .7)))

legend("bottom", c("Ignore Missing Data", "Input Mean to Missing Data"), lty=c(1,2),
       col=c("black",rgb(1, 0, 0, .7)))
```

# Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend"
```{r}
library(dplyr)
y <- mutate(y, newvar = weekdays(date))

weekday <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

y$newvar <- factor((y$newvar %in% weekday), levels=c(TRUE, FALSE),
                   labels=c("weekday", "weekend"))
str(y$newvar)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
y1 <- y
y1$newvar <- as.character(y1$newvar)

y1 <- aggregate(y1$steps, by=list(y1$interval, y$newvar), FUN=mean)

labels=c("weekday"="Weekday","weekend"="Weekend")

library(ggplot2)
g <- ggplot(y1, aes(Group.1, x))
g + geom_line() + 
        facet_wrap( ~Group.2, ncol=1, labeller=labeller(Group.2=labels)) +
        xlab("Interval") + ylab("Number of Steps")
```










